% References

\chapter*{References}
\addcontentsline{toc}{chapter}{References}

% Placeholder bibliography - replace with actual references
% This is a sample format for APA 7th edition citations

\begin{thebibliography}{9}

\bibitem{cavoukian2011privacy}
Cavoukian, A. (2011).
\textit{Privacy by design: The 7 foundational principles}.
Information and Privacy Commissioner of Ontario.

\bibitem{date2003database}
Date, C. J. (2003).
\textit{An introduction to database systems} (8th ed.).
Pearson Education.

\bibitem{diest2010graph}
Diest, R. van. (2010).
\textit{Graph theory with applications}.
Springer.

\bibitem{ghosh2007self}
Ghosh, S., et al. (2007).
Self-healing systems: A survey and perspective.
\textit{IEEE Transactions on Systems, Man, and Cybernetics}, 37(1), 1-15.

\bibitem{hevner2004design}
Hevner, A. R., March, S. T., Park, J., \& Ram, S. (2004).
Design science in information systems research.
\textit{MIS Quarterly}, 28(1), 75-105.

\bibitem{jennings1998agent}
Jennings, N. R., \& Wooldridge, M. J. (1998).
Applications of intelligent agents.
\textit{Agent Technology: Foundations, Applications, and Markets}, 3-26.

\bibitem{riquelme2018smonitor}
Riquelme, J. L., et al. (2018).
Smonitor: An IoT system for cattle monitoring.
\textit{Computers and Electronics in Agriculture}, 150, 168-176.

\bibitem{sandhu1996role}
Sandhu, R. S., Coyne, E. J., Feinstein, H. L., \& Youman, C. E. (1996).
Role-based access control models.
\textit{IEEE Computer}, 29(2), 38-47.

\bibitem{wolfert2017big}
Wolfert, J., Ge, L., Verdouw, C., \& Bogaardt, M. J. (2017).
Big data in smart farming – A review.
\textit{Agricultural Systems}, 153, 69-80.

\bibitem{wooldridge2009introduction}
Wooldridge, M. (2009).
\textit{An introduction to multiagent systems} (2nd ed.).
John Wiley \& Sons.

\end{thebibliography}
