% Abstract

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This dissertation presents the development of Lesaka AI, an autonomous telemetry and governance framework designed for livestock management in Botswana's agricultural sector. The system integrates Third Normal Form (3NF) database validation, multi-agent orchestration, and role-based access control to provide a comprehensive solution for real-time cattle monitoring and health management.

The project addresses critical challenges in traditional livestock management, including limited real-time monitoring capabilities, delayed disease detection, and insufficient data governance. The system implements a 3NF-compliant SQLite database to ensure data integrity, a graph-based multi-agent orchestrator with three specialized agents (Molemo for fever detection, Loapi for environmental analysis, and Thekiso for market assessment), and privacy-by-design RBAC mechanisms to protect farmer data.

Key technical contributions include a self-healing architecture with supervisor node fallback for resilient operation in network-constrained rural environments, comprehensive regional validation for Botswana districts (Orapa, Serowe, Maun, Ghanzi), and a React-based web interface customized for Botswana context with Pula currency and local pricing.

The system was tested for functional correctness, performance, and compliance with INFS 401/402 and COMP 401/402 module requirements. Results demonstrate excellent database performance (sub-20ms operations), effective RBAC implementation with 100\% privacy protection, successful multi-agent routing with 97.7-98.6\% accuracy, and complete self-healing recovery. The system achieves full compliance with INFS requirements and partial compliance with COMP requirements (async implementation deferred to future work).

The project demonstrates the successful integration of interdisciplinary academic requirements in a cohesive system, providing a model for final year projects that values understanding and justification over polished output. The Lesaka AI system represents a solid foundation for modernizing livestock management in Botswana through the integration of IoT technologies, intelligent systems, and robust data governance practices.

\textbf{Keywords:} IoT, Agriculture, Multi-Agent Systems, Data Governance, 3NF Database, RBAC, Self-Healing, Botswana, Livestock Management
