% Acknowledgements

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to express my sincere gratitude to all those who have contributed to the successful completion of this project.

First and foremost, I would like to thank my project supervisor, [Supervisor Name], for their invaluable guidance, support, and encouragement throughout the development of the Lesaka AI system. Their expertise and insights have been instrumental in shaping the direction and quality of this work.

I am grateful to the Botswana International University of Science and Technology (BIUST) Department of Computer Science for providing the academic framework and resources that made this project possible. The module coordinators for INFS 401/402 and COMP 401/402 deserve special recognition for their clear requirements and assessment criteria that guided the project development.

I would like to acknowledge the agricultural community in Botswana, particularly the farmers in Orapa, Serowe, Maun, and Ghanzi districts, whose real-world needs inspired this project. Their insights into livestock management challenges provided the practical context for this work.

My family deserves special thanks for their unwavering support and understanding throughout this academic journey. Their encouragement during challenging times has been a constant source of motivation.

I would also like to thank my colleagues and friends at BIUST for their constructive feedback, collaborative spirit, and moral support during the development process.

Finally, I acknowledge the broader open-source community whose tools, libraries, and documentation have made modern software development accessible and efficient. The Python, React, and Flask communities, in particular, have provided invaluable resources that accelerated the development of the Lesaka AI system.

This project represents not only an academic achievement but also a step toward modernizing agricultural practices in Botswana through technology. I hope that the work presented here will contribute to the continued advancement of the agricultural sector and serve as a foundation for future innovations.
