% References

\chapter*{References}
\addcontentsline{toc}{chapter}{References}

% Placeholder bibliography - replace with actual references
% This is a sample format for APA 7th edition citations

\begin{thebibliography}{9}

\bibitem{bag2019smart}
Bag, M., et al. (2019).
Smart farming: IoT-based agricultural monitoring system.
\textit{International Journal of Computer Applications}, 123(5), 45-52.

\bibitem{botswana_agriculture_2023}
Botswana Ministry of Agricultural Development and Food Security. (2023).
\textit{Agricultural sector report}.
Government of Botswana.

\bibitem{botswana_livestock_2022}
Botswana Ministry of Agricultural Development and Food Security. (2022).
\textit{Livestock production and economic impact report}.
Government of Botswana.

\bibitem{cavoukian2011privacy}
Cavoukian, A. (2011).
\textit{Privacy by design: The 7 foundational principles}.
Information and Privacy Commissioner of Ontario.

\bibitem{codd1970relational}
Codd, E. F. (1970).
A relational model of data for large shared data banks.
\textit{Communications of the ACM}, 13(6), 377-387.

\bibitem{damodaran2019data}
Damodaran, S. (2019).
\textit{Data governance: The definitive guide}.
Packt Publishing.

\bibitem{dara2018precision}
Dara, R., et al. (2018).
Precision agriculture in developing countries: Challenges and opportunities.
\textit{Computers and Electronics in Agriculture}, 155, 136-145.

\bibitem{date2003database}
Date, C. J. (2003).
\textit{An introduction to database systems} (8th ed.).
Pearson Education.

\bibitem{hevner2004design}
Hevner, A. R., March, S. T., Park, J., \& Ram, S. (2004).
Design science in information systems research.
\textit{MIS Quarterly}, 28(1), 75-105.

\bibitem{li2018database}
Li, X., et al. (2018).
Database design for IoT applications: A hybrid approach.
\textit{IEEE Transactions on Knowledge and Data Engineering}, 30(3), 456-468.

\bibitem{li2007t}
Li, N., Li, T., \& Venkatasubramanian, S. (2007).
t-closeness: Privacy beyond k-anonymity and l-diversity.
\textit{IEEE International Conference on Data Mining}, 106-115.

\bibitem{machanavajjhala2007diversity}
Machanavajjhala, A., et al. (2007).
l-diversity: Privacy beyond k-anonymity.
\textit{ACM Transactions on Knowledge Discovery from Data}, 1(1), 3.

\bibitem{navathe2020database}
Navathe, S. B., \& Elmasri, R. (2020).
\textit{Database system concepts} (7th ed.).
Pearson Education.

\bibitem{riquelme2018smonitor}
Riquelme, J. L., et al. (2018).
Smonitor: An IoT system for cattle monitoring.
\textit{Computers and Electronics in Agriculture}, 150, 168-176.

\bibitem{sweeney2002k}
Sweeney, L. (2002).
k-anonymity: A model for protecting privacy.
\textit{International Journal of Uncertainty, Fuzziness and Knowledge-Based Systems}, 10(5), 557-570.

\bibitem{wang1996beyond}
Wang, R. Y., \& Strong, D. M. (1996).
Beyond accuracy: What data quality means to data consumers.
\textit{Journal of Management Information Systems}, 12(4), 5-33.

\bibitem{wolfert2017big}
Wolfert, J., Ge, L., Verdouw, C., \& Bogaardt, M. J. (2017).
Big data in smart farming – A review.
\textit{Agricultural Systems}, 153, 69-80.

\end{thebibliography}
