% Chapter 2: Literature Review

\section{Introduction}

I started this literature review trying to figure out what's actually been done in agricultural data governance versus what people claim has been done. The research falls into four main areas: database normalization, data governance frameworks, privacy-by-design, and agricultural IoT systems. What I found was a lot of gaps between academic theory and what actually works in production environments, especially in places like Botswana where infrastructure constraints change everything.

\section{Database Normalization and Design}

Database normalization is one of those concepts that gets taught in every database course but gets applied inconsistently in real systems. Codd introduced the idea back in 1970 \cite{codd1970relational}, and it's still the standard way to eliminate redundancy and maintain integrity. But reading through the literature, I noticed that strict adherence to normalization rules often breaks down when you're dealing with high-volume IoT data.

\subsection{Third Normal Form (3NF)}

3NF builds on 1NF and 2NF by eliminating transitive dependencies \cite{date2003database}. Date lists the benefits clearly: no redundancy, no update anomalies, better integrity, easier maintenance. That all makes sense on paper.

The problem is that some researchers point out that strict 3NF can kill query performance \cite{navathe2020database}. This trade-off hit me when I was designing the Lesaka database. I initially wanted to normalize everything perfectly, but then I realized that for time-series telemetry data, the join overhead would be unacceptable on edge devices. I ended up keeping the core farmer and cattle data in strict 3NF but made some compromises on how the telemetry logs are structured.

\subsection{Database Design for IoT}

Li et al. \cite{li2018database} propose a hybrid approach that caught my attention: normalized tables for critical data combined with denormalized views for frequently accessed data. This is exactly the kind of practical compromise that makes sense in production. I considered this approach for Lesaka, but given that we're using SQLite on edge devices, maintaining separate views would add complexity without enough performance benefit. Instead, I focused on proper indexing to keep query times down while maintaining the 3NF structure.

Agricultural IoT has specific constraints that most database research doesn't account for:
- Time-series data comes in constantly, not in batches
- Network connectivity in rural Botswana is intermittent at best
- Edge devices have limited CPU and memory
- Farmers need real-time alerts, not batch reports

\section{Data Governance Frameworks}

Data governance sounds like an abstract enterprise concept until you actually try to build a system that handles real data. It's about managing availability, usability, integrity, and security \cite{damodaran2019data}. What I found reading through the literature is that most frameworks are designed for corporate environments with reliable infrastructure, not for rural agricultural settings where a power outage can take down your entire data pipeline.

\subsection{Data Quality Management}

Wang and Strong \cite{wang1996beyond} break data quality into four categories: intrinsic, contextual, representational, and accessibility. The framework is useful theoretically, but when I tried applying it to agricultural sensor data, I realized that "contextual" quality is the one that actually matters most. A temperature reading might be technically accurate (intrinsic quality), but if the sensor was sitting in direct sunlight for three hours, the context makes it useless for cattle health monitoring.

The data quality problems I actually ran into during testing:
- Sensors drifting after weeks in the field
- Cattle rubbing against collars causing intermittent readings
- Network drops during data transmission
- Missing data points when batteries die

\subsection{Data Integrity Constraints}

Elmasri and Navathe \cite{navathe2020database} emphasize that integrity constraints are what actually keep a database from turning into a mess. The four types are straightforward: entity integrity for unique IDs, referential integrity for valid relationships, domain integrity for valid values, and user-defined integrity for business rules.

When I implemented these in the Lesaka system, the referential integrity constraints were the ones that actually caught the most errors during testing. Farmers would try to register cattle with non-existent owner IDs, or telemetry would come in for cattle that had been removed from the system. The foreign key constraints prevented these orphaned records from ever making it into the database.

\section{Privacy-by-Design Principles}

Privacy-by-design is one of those concepts that sounds obvious in retrospect but gets ignored constantly in practice. Cavoukian \cite{cavoukian2011privacy} argues that privacy protections should be built into systems from the start, not bolted on as an afterthought. This is particularly relevant for agricultural data, where farmers have legitimate concerns about who can access their location data and commercial information.

\subsection{Core Principles}

Cavoukian's seven principles are comprehensive, but when I tried applying them to the Lesaka system, I found that some are more practical than others:

1. Proactive not reactive - This is the one that actually matters. Build privacy in before you have a breach.
2. Privacy as the default - Harder than it sounds when you're trying to make data accessible.
3. Privacy embedded into design - This required me to rethink the entire database schema.
4. Full functionality - Privacy shouldn't break the system's actual purpose.
5. End-to-end security - Standard security practices apply.
6. Visibility and transparency - Farmers need to know what's happening with their data.
7. Respect for user privacy - This is the baseline.

\subsection{Anonymization Techniques}

Sweeney's famous study \cite{sweeney2002k} showed that removing names isn't enough - you can re-identify 87\% of Americans using just zip code, gender, and date of birth. That finding scared me when I read it, because agricultural data has similar re-identification risks. A farmer's location combined with cattle count and breed can identify them even without their name.

The sophisticated techniques like k-anonymity, l-diversity, and t-closeness \cite{sweeney2002k, machanavajjhala2007diversity, li2007t} are mathematically sound but computationally expensive. For the Lesaka system, I went with a simpler approach: generate random 8-character owner IDs and separate personal data from telemetry data. It's not as robust as k-anonymity, but it's practical for edge deployment and provides reasonable privacy protection for the agricultural context.

\section{Agricultural IoT Systems}

Wolfert et al. \cite{wolfert2017big} review big data in smart farming, but most of the systems they describe are designed for developed countries with reliable infrastructure. When I tried applying these concepts to Botswana, I realized that the infrastructure assumptions don't hold. You can't assume constant connectivity, you can't assume farmers have smartphones, and you can't assume power is always available.

\subsection{Livestock Monitoring}

Riquelme et al. \cite{riquelme2018smonitor} built a cattle monitoring system using wireless sensor networks, and their results show the potential for early disease detection. What struck me about their work is that they focused almost entirely on the technical aspects of data collection - sensors, networks, transmission protocols. The data governance side is barely mentioned.

Bag et al. \cite{bag2019smart} point out this exact problem: agricultural IoT systems collect massive amounts of data but don't have standardized governance frameworks. This is the gap I'm trying to address with Lesaka. The system needs to handle data properly, not just collect it.

\subsection{Challenges in Developing Regions}

Dara et al. \cite{dara2018precision} emphasize that solutions need to be context-aware, but reading their work made me realize how much context actually matters. The standard challenges are obvious:

- Internet connectivity is intermittent or non-existent in rural areas
- Edge devices have limited CPU, memory, and storage
- Technical infrastructure is unreliable
- Farmers may not have high digital literacy

But the real challenge I discovered during testing is cultural. Farmers trust their traditional methods, and convincing them to adopt sensor-based monitoring requires building trust through demonstrated reliability, not just technical superiority. A system that works perfectly in a lab can fail completely in the field if farmers don't trust it.

\section{Research Gaps}

Going through the literature, I found four specific gaps that justify the Lesaka project:

\subsection{Gap 1: Data Governance in Agricultural IoT}

Agricultural IoT systems get plenty of attention in the literature, but almost all of it focuses on data collection and analysis \cite{bag2019smart}. The governance side - how to validate the data, who can access it, how to protect it - is barely mentioned. This is a problem because collecting data without proper governance just creates a bigger mess. You end up with inaccurate data, privacy violations, and systems that can't scale.

\subsection{Gap 2: 3NF in Edge Computing}

Li et al. \cite{li2018database} acknowledge that normalization-performance trade-offs need more study, but there's almost no research on how 3NF actually performs on resource-constrained edge devices. My testing showed that strict 3NF is workable if you use proper indexing, but this needs more systematic study. The literature assumes you have a powerful database server, not a Raspberry Pi running SQLite in the middle of a cattle post.

\subsection{Gap 3: Privacy-by-Design in Agriculture}

Privacy-by-design is well-studied in healthcare and finance \cite{cavoukian2011privacy}, but agricultural privacy concerns are different. Farmers aren't worried about medical records - they're worried about competitors knowing their cattle counts, location data revealing their grazing patterns, and commercial information being exposed. The existing privacy frameworks don't account for these agricultural-specific concerns.

\subsection{Gap 4: Botswana Context}

This is the most obvious gap. Almost all agricultural IoT research assumes developed-country infrastructure \cite{botswana_agriculture_2023}. There's minimal research on systems designed specifically for Botswana's constraints: intermittent connectivity, limited technical infrastructure, and the cultural context of traditional farming practices. A system that works in the Netherlands won't necessarily work in rural Botswana.

\section{Comparison of Existing Systems}

Table \ref{tab:comparison} shows how existing systems stack up against the Lesaka framework in terms of data governance features.

\begin{table}[H]
\centering
\caption{Comparison of Agricultural IoT Systems}
\label{tab:comparison}
\begin{tabular}{|l|c|c|c|c|}
\hline
\textbf{System} & \textbf{3NF DB} & \textbf{RBAC} & \textbf{Privacy} & \textbf{Audit} \\
\hline
Riquelme et al. \cite{riquelme2018smonitor} & No & No & No & No \\
Bag et al. \cite{bag2019smart} & Partial & No & No & Yes \\
Dara et al. \cite{dara2018precision} & No & Yes & No & Yes \\
\textbf{Lesaka AI (Proposed)} & \textbf{Yes} & \textbf{Yes} & \textbf{Yes} & \textbf{Yes} \\
\hline
\end{tabular}
\end{table}

The table makes it obvious that existing systems are missing comprehensive data governance. Riquelme's system has no governance at all - it's purely data collection. Bag and Dara have some features but are incomplete. Lesaka is the only one that combines all four: proper database design, access control, privacy protection, and audit logging. This combination is what makes it different from existing work.

\section{Conclusion}

This literature review gave me the theoretical foundation I needed, but more importantly, it showed me where the gaps are. Database normalization is well-understood but not well-applied in IoT contexts. Data governance frameworks exist but aren't designed for agricultural settings. Privacy-by-design is studied but not adapted for farmer concerns. And almost nothing accounts for Botswana's specific infrastructure and cultural constraints.

The next chapter explains how I turned these theoretical concepts into an actual working system, including the methodology I used and the practical decisions I made along the way.
