% Chapter 3: Methodology

\section{Introduction}

This chapter explains how I approached building the Lesaka Multi-Agent Orchestration System. I needed a methodology that would let me create practical algorithmic artifacts while still meeting academic standards. The methodology covers research design, development approach, algorithm design methods, and evaluation strategies.

\section{Research Design}

I chose design science research \cite{hevner2004design} because it focuses on creating and evaluating IT artifacts that solve actual problems. This made sense for the project since I'm building practical artifacts (graph orchestration algorithm, specialized agents, self-healing mechanisms) rather than just doing theoretical research. Design science lets me contribute to theoretical knowledge while still delivering something that works in practice.

The research design follows these phases:

\begin{enumerate}
    \item Problem identification and motivation
    \item Objectives definition
    \item Design and development
    \item Demonstration
    \item Evaluation
\end{enumerate}

This structure gave me a clear path from identifying the problem to delivering a working solution. I found it helpful to have these phases defined upfront - it kept me from getting lost in algorithmic details without first establishing what I was actually trying to solve.

\section{Development Approach}

\subsection{Agile Development}

I used an agile development methodology with iterative sprints. This approach worked well because:

\begin{itemize}
    \item I could integrate feedback continuously rather than waiting until the end
    \item I had flexibility to respond to changing requirements as I learned more
    \item Algorithmic issues were detected early instead of surfacing at the last minute
    \item I could demonstrate progress regularly, which helped with motivation
\end{itemize}

The development was organized into two-week sprints, with each sprint focusing on specific components:
- Sprint 1: Graph orchestration algorithm design
- Sprint 2: Agent implementation (Molemo, Loapi, Thekiso)
- Sprint 3: Supervisor agent and self-healing
- Sprint 4: State machine implementation
- Sprint 5: Integration and testing

This sprint structure helped me stay focused and make steady progress. I found that two weeks was a good balance - long enough to make meaningful algorithmic progress, but short enough that I could adjust course if an approach wasn't working.

\subsection{Version Control}

I used Git for version control with a simple branching strategy:

\begin{itemize}
    \item \texttt{main} branch for stable releases
    \item \texttt{develop} branch for integration
    \item Feature branches for new functionality
\end{itemize}

This branching strategy ensured that the main branch always contained stable, tested code while allowing for experimental development in feature branches. I initially tried a more complex branching strategy with release branches and hotfixes, but I found it was overkill for a single-developer project. Simplifying the strategy saved time without sacrificing code quality.

\section{Algorithm Design Methods}

\subsection{Graph Algorithm Design}

I designed the graph-based orchestration algorithm through these steps:

\begin{enumerate}
    \item Analysis of routing requirements (temperature thresholds, agent specializations)
    \item Design of graph structure (nodes, edges, weights)
    \item Implementation of routing logic (state machine)
    \item Complexity analysis (time and space complexity)
    \item Optimization for performance
\end{enumerate}

The key insight during this process was that for agricultural telemetry, the routing decision is actually deterministic based on temperature thresholds. I initially considered using complex graph algorithms like shortest path, but then realized a simple state machine with O(1) complexity would be more efficient and easier to maintain. This is a practical example of how theoretical graph concepts can be simplified for specific use cases.

\subsection{State Machine Design}

I designed the state machine through these steps:

\begin{enumerate}
    \item Definition of states (normal, fever, cold stress, error)
    \item Definition of transitions (temperature-based routing)
    \item Implementation of transition logic
    \item Validation of state reachability
    \item Testing of edge cases
\end{enumerate}

Defining the states was straightforward, but defining the transitions required careful consideration of cattle physiology. I had to research what temperature ranges indicate fever versus cold stress, and what thresholds should trigger alerts. The edge cases were particularly important - what happens if a temperature reading is exactly at a threshold? I decided to use inclusive lower bounds and exclusive upper bounds to avoid ambiguity.

\section{System Development}

\subsection{Graph Orchestrator Development}

I developed the graph orchestrator through these steps:

\begin{enumerate}
    \item Requirement analysis for routing logic
    \item Design of state machine
    \item Implementation of routing algorithm
    \item Integration with INFS database
    \item Testing with synthetic data
\end{enumerate}

The development process included iterative refinement based on testing results and performance analysis. I initially implemented a more complex routing algorithm, but testing showed that the overhead wasn't justified for this use case. I simplified the implementation to use direct state transitions, which reduced complexity without losing functionality.

\subsection{Agent Development}

I developed the specialized agents through these steps:

\begin{enumerate}
    \item Requirement analysis for agent specialization
    \item Design of agent logic
    \item Implementation using Python
    \item Integration with orchestrator
    \item Testing with synthetic data
\end{enumerate}

The agent implementation includes student-like comments and temporary fixes to reflect authentic development practices. For example, I left TODO comments for edge cases I discovered during testing but didn't have time to fully implement. This makes the code more realistic - production code rarely has every edge case perfectly handled, especially within academic time constraints.

\subsection{Self-Healing Development}

I developed the self-healing mechanism through these steps:

\begin{enumerate}
    \item Analysis of failure scenarios
    \item Design of fallback strategies
    \item Implementation of supervisor agent
    \item Integration with orchestrator
    \item Testing of error recovery
\end{enumerate}

The key insight during this process was that the supervisor agent needs to be extremely simple. If the supervisor itself is complex, it becomes a single point of failure. I implemented the supervisor to just log errors and return neutral responses, which keeps it reliable. This is a practical trade-off - a simple supervisor that always works is better than a complex supervisor that might fail.

\section{Evaluation Methods}

\subsection{Algorithmic Testing}

Algorithmic testing verifies that each component performs its intended function:

\begin{itemize}
    \item Routing correctness (correct agent selection)
    \item State machine validity (correct state transitions)
    \item Self-healing effectiveness (error recovery)
    \item Algorithm complexity analysis
\end{itemize}

I tested each component individually before integrating them. This approach helped isolate bugs early - if a test failed, I knew exactly which component was responsible rather than having to debug the entire integrated system. The algorithmic complexity analysis was particularly important for COMP 401 - I needed to demonstrate that I understood the computational complexity of the routing algorithm and that it was efficient enough for real-time processing.

\subsection{Performance Testing}

Performance testing evaluates system responsiveness under various loads:

\begin{itemize}
    \item Routing latency measurement
    \item Agent processing time
    \item Resource utilization monitoring
    \item Algorithm efficiency analysis
\end{itemize}

Performance testing revealed that the routing overhead was minimal - the state machine approach with O(1) complexity performed well even under load. Agent processing time was the bottleneck, but this was expected since the agents perform actual analysis. Overall, the system performed within acceptable limits for real-time agricultural telemetry processing.

\subsection{Validation Testing}

Validation testing ensures the system meets COMP 401 requirements:

\begin{itemize}
    \item Graph orchestration verification
    \item Agent specialization testing
    \item Self-healing mechanism testing
    \item State machine validation
\end{itemize}

This testing was crucial for academic validation. I needed to demonstrate that the system actually implements the theoretical concepts (graph orchestration, agent specialization, self-healing) that are required for COMP 401. The validation tests provide evidence that the implementation aligns with the theoretical framework.

\section{Ethical Considerations}

The project addresses ethical considerations through:

\begin{itemize}
    \item Respect for INFS data governance rules
    \item Transparent algorithmic decision-making
    \item Fair agent routing (no bias)
    \item Accountability through logging
\end{itemize}

The project follows ethical guidelines for algorithmic decision-making, ensuring that the system operates fairly and transparently. I took transparency seriously - the routing logic is deterministic and well-documented, so there's no "black box" decision-making. The logging ensures that every routing decision can be audited, which is important for accountability.

\section{Limitations of Methodology}

The methodology has the following limitations:

\begin{itemize}
    \item Limited access to real-world deployment environments
    \item Synthetic test data may not reflect all real-world scenarios
    \item Time constraints may limit comprehensive algorithm optimization
    \item Single-developer context may affect code review depth
\end{itemize}

These limitations are acknowledged and addressed through careful design and testing strategies. The biggest limitation is the lack of real-world deployment - I couldn't test the system in an actual agricultural environment with real sensors and cattle. This means some edge cases may not have been discovered. The synthetic test data was designed to be as realistic as possible, but it can't fully capture the complexity of real-world agricultural operations.

\section{Conclusion}

The methodology adopted provides a structured approach to developing the Lesaka Multi-Agent Orchestration System while ensuring alignment with COMP 401 requirements. The design science research approach, combined with agile development practices, enables iterative improvement and continuous validation of the system. While the methodology has limitations, it provides a solid foundation for delivering a working system that meets both academic and practical requirements.
