% References

\chapter*{References}
\addcontentsline{toc}{chapter}{References}

% Placeholder bibliography - replace with actual references
% This is a sample format for APA 7th edition citations

\begin{thebibliography}{9}

\bibitem{dara2018precision}
Dara, R., et al. (2018).
Precision agriculture in developing countries: Challenges and opportunities.
\textit{Computers and Electronics in Agriculture}, 155, 136-145.

\bibitem{diest2010graph}
Diest, R. (2010).
\textit{Graph theory with applications} (5th ed.).
Springer.

\bibitem{ghosh2007self}
Ghosh, S., et al. (2007).
Self-healing systems: A survey and taxonomy.
\textit{ACM Computing Surveys}, 39(3), 7.

\bibitem{hevner2004design}
Hevner, A. R., March, S. T., Park, J., \& Ram, S. (2004).
Design science in information systems research.
\textit{MIS Quarterly}, 28(1), 75-105.

\bibitem{jennings1998agent}
Jennings, N. R., \& Wooldridge, M. (1998).
Applications of intelligent agents.
\textit{Agent Technology: Foundations, Applications, and Markets}, 3-28.

\bibitem{kraus1997negotiation}
Kraus, S. (1997).
Negotiation and cooperation in multi-agent environments.
\textit{Artificial Intelligence}, 94(1-2), 79-97.

\bibitem{luck2003agent}
Luck, M., McBurney, P., \& Preist, C. (2003).
A manifesto for agent technology: Towards next generation computing.
\textit{Autonomous Agents and Multi-Agent Systems}, 7(1-2), 119-125.

\bibitem{parunak2001practical}
Parunak, H. V. D., Savic, R., \& Riolo, R. L. (2001).
Agent-based modeling of industrial ecosystems.
\textit{Proceedings of the 2001 Winter Simulation Conference}, 447-452.

\bibitem{riquelme2018smonitor}
Riquelme, J. L., et al. (2018).
Smonitor: An IoT system for cattle monitoring.
\textit{Computers and Electronics in Agriculture}, 150, 168-176.

\bibitem{sycara1998multiagent}
Sycara, K. P. (1998).
Multiagent systems.
\textit{AI Magazine}, 19(2), 79-92.

\bibitem{wooldridge2009introduction}
Wooldridge, M. (2009).
\textit{An introduction to multiagent systems} (2nd ed.).
John Wiley \& Sons.

\end{thebibliography}
